---
title: "R Notebook"
output: html_notebook
---



de la Mata-Espinosa P, Bosque-Sendra JM, Bro R, Cuadros-Rodriguez L, Discriminating olive and non-olive oils using HPLC-CAD and chemometrics, Analytical and Bioanalytical Chemistry, 2011a, 399, 2083-2092.
de la Mata-Espinosa P, Bosque-Sendra JM, Bro R, Cuadros-Rodriguez L, Olive oil quantification of edible vegetable oil blends using triacylglycerols chromatographic fingerprints and chemometric tools, Talanta, 2011b, 85, 177-182.

```{r}
data()
tmp <- readMat("/Users/katiedaisey/Downloads/HPLCforweb.mat")
data <- tmp$HPLCforweb$data
label <- tmp$HPLCforweb$label[[1]][[1]]
class <- tmp$HPLCforweb$class[[1]][[1]]
matplot(t(data), type = "l",
        xlab = "Time", ylab = "AU")
```

```{r}
cltrain <- class[c(T,F)]
cltest <- class[c(F,T)]
train <- data[c(T,F),]
test <- data[c(F,T),]

matplot(t(data), type = "l", lty = 1, col = rep(c(1,2), times = 60))
```




```{r}
matplot(t(data), type = "l", lty = 1, col = class + 1,
        xlab = "Time", ylab = "AU", main = "Olive Oils")
```






```{r}
?knn
knn1 <- class:::knn(train, test, cl = cltrain, k = 1)
plot(1:60, knn1, col = cltest, pch = 20)
```



```{r}
scale_train <- scale(train, center = T, scale = T)
scale_test <- scale(test, center = attributes(scale_train)$`scaled:center`, scale = attributes(scale_train)$`scaled:scale`)


matplot(t(scale_train), type = "l", lty = 1, col = cltrain,
        xlab = "Time", ylab = "AU, Scaled, Centered")
```

```{r}
knn2 <- class:::knn(scale_train, scale_test, cltrain, k = 1)
plot(1:60, knn2, col = cltest, pch = 20,
     xlab = "Sample No.", ylab = "Predicted Class",
     main = "knn, scaled")
```


```{r}

```




```{r}
hmlcdir <- "/Users/katiedaisey/Desktop/hmlc"

library(class)
library(caret)
library(ggplot2)
library(pls)
library(randomForest)
library(rpart)
library(devtools)
install(hmlcdir)
library(hmlc)
```



```{r}
gen_hmlc_tree <- function(classes, data, dist_method = "euclidean", hclust_method = "ward.D2") {
  
  tmp_df <- data.frame(Class = classes)
  tmp_df$Data <- as.matrix(data)
  centers <- aggregate(Data~Class, data = tmp_df, FUN = mean)
  
  h1 <- stats:::hclust(stats:::dist(centers, method = dist_method), method = hclust_method)
  h1 <- hmlc:::hclust_to_hltree(h1)
  
  return(h1)
}
# make tree
h1 <- gen_hmlc_tree(cltrain, train)

flatten <- function(results, leaf_nodes) {
  results <- results[,leaf_nodes]
  
  out <- rep(0, length(results[,1]))
  
  for (i in leaf_nodes) {
    out[results[,i] == 1] <- i
  }
  
  return(out)
}  
library(hmlc)
labels <- create_labels(cltrain, hltree = h1, depth = "full")
plot(h1)
```


```{r}
hmlc1 <- hmlc:::hsc.lcn(data = train, test = test, labels = labels, hltree = h1, policy = "siblings", classifier = "randomForest", k = 1)
a <- flatten(hmlc1, 1:3)

plot(1:60, a, pch = 20, col = cltest, 
     xlab = "Sample No.", ylab = "Predicted Class")
```





